---
title: "Assignment 2"
author: "khushboo.satpute"
date: "13/11/2019"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Q1. Create R Markdown


## Q2 Load in the data as an object called DublinAirport
```{r}
DublinAirport<-read.csv("mly532.csv", header=TRUE, sep=",", skip = 19)
str(DublinAirport)
```
## Q3. Transform the column months.
```{r}
DublinAirport$month<-factor(DublinAirport$month, labels = month.name)
head(DublinAirport)
```
## Q4. Using  aggregate function to compute which month has on average the highest and the lowest Precipitation Amount.
```{r}
Agg_dub_rain = aggregate(DublinAirport$rain,by = list(DublinAirport$month) , FUN = mean,na.rm= TRUE)
message("The Minimum Average Precipitation Amount month is")
as.character(Agg_dub_rain$Group.1[Agg_dub_rain$x == max(Agg_dub_rain$x)])
message("The Minimum average Precipitation amount Month is")
as.character(Agg_dub_rain$Group.1[Agg_dub_rain$x == min(Agg_dub_rain$x)])

```
## Q5.Adding column which contains a factor indicating the season:
```{r}
library(dplyr)
DublinAirport<-mutate(DublinAirport,season = case_when( month %in% c("December","January","February") ~ "Winter",
                           month %in% c("March","April","May")   ~ "Spring",
                           month %in% c("June","July","August") ~ "Summer",
                           month %in% c("September","October","November") ~ "Autumn",
                          ))
DublinAirport$season<-factor(DublinAirport$season,labels=c("Winter","Spring","Summer","Autumn"))
head(DublinAirport)
```
## Q6. Assign to the DublinAiport object the classes WeatherData and data.frame
```{r}
class(DublinAirport)<- c('weatherdata','data.frame')
class(DublinAirport)
```

## Q7. The S3 summary method for an object of class WeatherData 
```{r}
Summary.WeatherData <- function(val){
  val %>%
    group_by(season) %>%
    select(rain, maxtp, mintp, maxgt) %>%
    summarise_all(funs(mean, max, sd), na.rm = TRUE)
}
Summary.WeatherData(DublinAirport)
```
## Q8. S3 plot method for the class WeatherData that produces the following plots.
#(1) Plot of the monthly Air Temperature (C) (maxtp, mintp).
#(2) Plot of the Precipitation Amount (mm) (rain).
#(3) Plot of the Highest Gust (knot) (maxgt).
```{r}
library(gridExtra)
library(ggplot2)
plot.WeatherData <- function(plotdata, start_year=2015, end_year=2018,
                             plot1 = TRUE, plot2 = TRUE, plot3 = TRUE ) 
  {
  
  df_sub <- subset(DublinAirport,DublinAirport$year >= 2015 & DublinAirport$year <= 2018)
  data <- data.frame(Timeseries = with(df_sub,
                                     sprintf("%d-%02d", df_sub$year, df_sub$month)),
                     mintp = df_sub$mintp,maxtp = df_sub$maxtp,rain = df_sub$rain,
                     maxgt = df_sub$maxgt)

  plt <- ggplot(data, aes(x= Timeseries)) +
    theme(axis.text.x = element_text(angle = 90, hjust = 1), legend.position="right") +
    labs(y = "Values") + scale_color_manual(values=c('#999999','#E69F00'))
  
  if(plot1 == TRUE & plot2 == FALSE & plot3 == FALSE ){
    plt <- plt + 
      geom_line(aes(y = mintp, group = 1, color = "Mintp")) + 
      geom_line(aes(y = maxtp, group = 1, color = "Maxtp")) +
        ggtitle("Monthly Air Temperature (C) (maxtp, mintp)")+theme(plot.title = element_text(hjust = 0.5))
    plt
  }
  else if(plot1 == FALSE & plot2 == TRUE & plot3 == FALSE ){
    plt <- plt + 
      geom_line(aes(y = rain, group = 1, color = "Rain")) +
      ggtitle("Monthly Rain in mm")+theme(plot.title = element_text(hjust = 0.5))
    plt
  }
  else if(plot1 == FALSE & plot2 == FALSE & plot3 == TRUE ){
    plt <- plt + 
      geom_line(aes(y = maxgt, group = 1, color = "Gust")) +
      ggtitle("Monthly Gust")+theme(plot.title = element_text(hjust = 0.5))
    plt
  }
  else if(plot1 == TRUE & plot2 == TRUE & plot3 == FALSE ){
    plt_1 <- plt + 
      geom_line(aes(y = mintp, group = 1, color = "Mintp")) + 
      geom_line(aes(y = maxtp, group = 1, color = "Maxtp"))  +
      labs( title = 'Air Temperature')
    plt_2 <- plt +
      geom_line(aes(y = maxgt, group = 1, color = "Rain")) +
      labs( title = 'Rain')
    grid.arrange(plt_1,plt_2)
  }
  else if(plot1 == TRUE & plot2 == FALSE & plot3 == TRUE ){
    plt_1 <- plt + 
      geom_line(aes(y = mintp, group = 1, color = "Mintp")) + 
      geom_line(aes(y = maxtp, group = 1, color = "Maxtp")) +
      labs( title = 'Air Temperature')
    plt_2 <- plt +
      geom_line(aes(y = maxgt, group = 1, color = "Gust")) +
      labs( title = 'Gust')
    grid.arrange(plt_1,plt_2)
  }
  else if(plot1 == FALSE & plot2 == TRUE & plot3 == TRUE ){
    plt_1 <- plt + 
      geom_line(aes(y = rain, group = 1, color = "rain")) +
      labs( title = 'Rain')
    plt_2 <- plt +
      geom_line(aes(y = maxgt, group = 1, color = "Gust")) +
      labs( title = 'Gust')
    grid.arrange(plt_1,plt_2)
  }else{
    plt_1 <- plt + 
      geom_line(aes(y = mintp, group = 1, color = "Mintp")) + 
      geom_line(aes(y = maxtp, group = 1, color = "Maxtp")) +
      labs( title = 'Air Temperature')
    plt_2 <- plt +
      geom_line(aes(y = rain, group = 1, color = "Rain")) +
      labs( title = 'Rain')
    plt_3 <- plt +
      geom_line(aes(y = maxgt, group = 1, color = "Gust")) +
      labs( title = 'Gust')
    grid.arrange(plt_1,plt_2,plt_3)
  }
}
# Plot 1
plot.WeatherData(DublinAirport, 2015, 2017, TRUE, FALSE, FALSE)
# Plot 2
plot.WeatherData(DublinAirport, 1990, 2000, FALSE, TRUE, TRUE)
# Plot 3
plot.WeatherData(DublinAirport )
```

